---
title: "Chick Weight Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(knitr)
```

```{r readin,echo=FALSE}
chicks <- read.csv('chickwts.txt', header = TRUE, sep = ",")
```

We have data which details the **weight** of different chicken bracketted by the type of **feed* they ate.  The whole of the data is summarized below:

```{r summary, echo=FALSE}
summary(chicks)
```

We summarize the data for **feed type** cassien below:

```{r cassein summary, echo=FALSE}
summary(chicks[which(chicks$feed=="cassein"),])
```

We summarize the data for **feed type** horsebean below:

```{r horsebean summary, echo=FALSE}
summary(chicks[which(chicks$feed=="horsebean"),])
```

We summarize the data for **feed type** linseed below:

```{r linseed summary, echo=FALSE}
summary(chicks[which(chicks$feed=="linseed"),])
```

We summarize the data for **feed type** meatmeal below:

```{r meatmeal summary, echo=FALSE}
summary(chicks[which(chicks$feed=="meatmeal"),])
```

We summarize the data for **feed type** soybean below:

```{r soybean summary, echo=FALSE}
summary(chicks[which(chicks$feed=="soybean"),])
```

We summarize the data for **feed type** sunflower below:

```{r sunflower summary, echo=FALSE}
summary(chicks[which(chicks$feed=="sunflower"),])
```

We now summarize the entirey of the data in a graph below:

```{r chicks, echo=FALSE}
d = ggplot(data = chicks)
d + geom_bar(aes(x=as.factor(feed),y=weight),stat="summary",
             fun.y="mean",fill="blue",color="black") + theme_classic() +
            xlab("Type of Feed")+ylab("Average Weight ")
```

To investigate whether feeding chickens different **feed** may have had an affect on their **weight**, we test whether the **weights** of chickens fed sunflower and soybeans are statistically different.  To do so, we first construct null and alternate hypotheses.

*Null Hypothesis: The* **weight** *of chickens fed sunflower and soybeans are not significantly different.*

*Alternate Hypothesis: The* **weight** *of chickens fed sunflower and soybeans are significantly different.*

We performed likelihood ratio test to compare the **weights** of chickens fed sunflower and soybeans and arrived at a p-value of approximately 0.000233458.  This is small enough to reject the null hypothesis, so we project that that the **weights** of chickens fed soybeans and sunflowers are significantly different.