---
title: "Exe12_Question2"
author: "SZD"
date: "November 17, 2017"
output: html_document
---

#Exe12 Question 2

## Military Time

For this question we need to find a regex that can discover military time between 12:01 and 23:59. Naturally, there are some restraints that exist in creating this regex due to the unusual nature of military time and the restrictions on which times the problem calls for. For examples, if we divide the regex into the five characters (numberic, nu the first character is a 2 (eg. 22:10, 23:45, etc), the second character can only be between 0-3, whilst if the first character is 1 (eg. 12:13, 18:40, etc.), then the second character can only be between 2-9. As such, what needs to be done to solve for this problem is to break down the times we want into if else statements, and within each if else statement write down the regex for a specifici military time. The code is shown below.

```{r}
library(stringr)

all_hours <- c("07:30", "10:15", "10:90", "12:01","15:30","15:80", "20:01","23:59","24:00","25:01", "23:591", "123:45", "18:00")

all_hours_answers <- vector(length = length(all_hours))

for(i in 1:length(all_hours)){
  if(str_detect(all_hours[i],"^1")==TRUE){ #1
    if(str_detect(all_hours[i],"^12:0")){ #2
      all_hours_answers[i] <- str_detect(all_hours[i],"^12:0[1-9]$") #3
    }
    else{
      all_hours_answers[i] <- str_detect(all_hours[i],"^1[3-9]:[0-5][0-9]$") #4
    }
  }
  else{
    all_hours_answers[i] <- str_detect(all_hours[i],"^2[0-3]:[0-5][0-9]$") #5
  }
  if(all_hours_answers[i]==TRUE){
    print(all_hours[i])
  }
  else{}
}

```

Briefly, in the above code, all_hours is a set of made up and real military times with some correct and some incorrect answers, whilst all_hours_answers is a vector where the entries are TRUE or FALSE depending on whether the corresponding cell in all_hours fit the criteria of the problem or not. The #1 if statement filters out any entries starting with a 1, and #2 if statement further narrows that down to any cases between 12:01 and 12:09, since this is the only case where a 0 is not allowed in the last place, which is checked by #3. #4 checks all other military times starting with 1, and #5 checks for all military times starting with 2. In each case we use ^ to specify beginning and $ to specify end so as to pickup the 5 digits only that is standard for military time. Finally, the for loop prints which times in our vector came back true.


## Species Names

Although similar to the above case, this one is much simpler. The requirements for this one are that the first character be a captial letter, the second character a period, the third character be a space, and the fourth character onwards be only lower case letters. The regex below achieves that by allowing lower case letters to be repeated indefinitely for the species name.

```{r}

spec_names <- c("H. sapien", "H. Sapien", "Ho. sapien", "H sapien", "H. sapieN", "hO. sapien")
spec_names[which(str_detect(spec_names,"^[[:upper:]].[ ][[:lower:]]+$"))]

```


## SSIDs

Again, similar to the military time question, SSIDs have a unique pattern of three numeric, hyphen, two numeric, hypen, four numeric, end. The below acheives that by using {} to specify the number of repeats and ^ and $ to specify the beginning and the end of the regex respectively.

```{r}

ssids <- c("123-45-6789","1234-56-7890", "12-34-5678", "123-456-7890", "123-4-5678", "123-45-67890", "123-45-678")
ssids[which(str_detect(ssids,"^[[:digit:]]{3}-[[:digit:]]{2}-[[:digit:]]{4}$"))]


```
