---
title: "Exercise 12"
author: "Michelle"
date: "November 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regular Expressions

Regular expressions (regex) are used to match desired strings in text or files from those that are unwanted by using code to demarcate significant portions of the string.

### Times after noon, but before midnight when reported in 24-hour or "military" format (e.g. 15:30)

Time range = 12:00 - 23:59

```{r}
"[12-23]:[00-59]" 
```

[ ] denotes matching to a single character

### Genus species names expressed in the format G. species (e.g. H. sapien)

```{r}
"[A-Z](.)( )[a-z]+"
```

[A-Z] is a single capitalized letter

(.) requires the "." character"

( ) requires a space after "."

[a-z]+ allows for one or more lowercase letter

### Social Security Numbers (e.g. 389-05-4771)

```{r}
"[0-9]{3}-[0-9]{2}-[0-9]{4}"
```

[0-9] searches for a single number from 0-9

{#} matches a digit where # is the number of characters in the square brackets it will match.
