---
title: "Problem2"
author: "Shane Davitt, Kevin Fick, and Logan Arnold"
date: "November 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction
A file (ExamplesforRegex.txt) contains all of the examples for problem 2, including data that should and shoul not be selected by the varous regex expressions. It is printed below:
```{r, echo=TRUE, eval=TRUE}
exampleData=scan(file="ExamplesforRegex.txt", what = "character")
exampleData
```
##a)
Regex1 <- "[12][0-9]:[0-5][0-9]"
```{r, echo=TRUE,eval=TRUE}
#This setup reused for later parts as well
library(regexr)

aOutput=grep("[12][0-9]:[0-5][0-9]", exampleData, value = TRUE)
aOutput
```
Comparing the results to the examples file reveals that only the desired times were selected.

***

##b)
Regex2 <- "[A-Z][.]\s[a-z]+($|\s)"
```{r, echo=TRUE,eval=TRUE}
bOutput=grep("[A-Z][.]\\s[a-z]+($|\\s)", exampleData, value = TRUE)
bOutput
```
Comparing the results to the examples file reveals that only the desired names were selected.

***

##c)
```{r, echo=TRUE,eval=TRUE}
cOutput=grep("^[0-9]{3}-[0-9]{2}-[0-9]{4}", exampleData, value = TRUE)
cOutput
```
Comparing the results to the examples file reveals that only the desired social security numbers were selected.