---
title: "Tutorial 12 - Problem 2"
author: "Chelsea Weibel"
date: "November 17, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
```

### a. Time after noon, but before midnight (12:00-23:59)

```{r}
time=c('1500','23:59', '12:40', '01:40', 'Cow')

print(str_detect(time,'(1[2-9]|2[0-3]):[0-5][0-9]'))

```

This expression will select all times reported in the afternoon (from 12:00-23:59) because the expression reads as follows:  
* 1[2-9] | 2[0-3] = there can be a 1 in the first position if there is a 2-9 in the second position (i.e. hours 12:00-19:00) OR there can be a 2 in the first position if there is a 0-3 in the second position (i.e. hours 20:00-23:00)  
* : = there will be a colon in the third position  
* [0-5] = there will be a 0-5 in the fourth (minute) position  
* [0-9] = there will be any digit in the fifth (seconds) position  


### b. Genus species names expressed in the format G. species

```{r}
species=c('P. cynocephalus','Dog', 'C. lupus', 'H. sapiens', 'Cow')

print(str_detect(species,'([A-Z][.] [a-z]+)'))

```

This expression will select only genus species names in the format G. species because it reads as follows:  
* [A-Z] = the first letter must be a capital  
* [.] = a period must follow in the second position  
* Followed by a space in the third position  
* [a-z]+ = there will be at least one lower case letter followed by any number of lowercase letters  

### c. Social security numbers (###-##-####)

```{r}
SSN=c('111-11-1111','555-555-4285', '919-19-1919', '56', 'Cow')

print(str_detect(SSN,'[0-9]{3}-[0-9]{2}-[0-9]{4}'))

```

This expression will select for all social security numbers becuase it reads as follows:  
* [0-9]{3} = there must be three of any number  
* - = followed by a dash  
* [0-9]{2} = then two of any number  
* - = followed by another dash  
* [0-9]{4} = then four of any number  